### 2025-08-19 response.txt
ready for revisions

Response to first referee report.

I thank the referee for their concise and insightful review of the paper.  The
main issue raised is that this paper relies heavily on the companion paper
analysing LSST constraints on asteroid masses from mutual encounters.  That
paper was submitted to PSJ some weeks before this one, and has since been
refereed, revised, accepted, and published in the journal, which I hope will
assuage this worry.

One of the revisions made to that paper addressed an issue also raised by this
referee, which was the rather high value of albedo assumed for the main-belt
asteroid population.  To address this concern for the other paper, we changed
our conversion from (well-observed) H values to (physically relevant) mass
substantially,  Instead of choosing a representative value of  both density and
albedo of the population, we surveyed the literature for individual main-belt
asteroids (MBAs) with mass determinations of any useful precision, then examined
the ratio $M/ 10^{-0.6H}$ directly.  The number of such bodies is actually
larger than the number for which both density and albedo have been determined
to similar precision, so is a simpler and slightly more accurate means of deriving
the "mass-to-light" conversion.  See Figure 1 and Section 2.1 of the published
paper.

The adopted result was that the distribution of mass-to-light
ratios of this (incomplete, heterogeneous, almost certainly biased) population
could be crudely characterized by taking a "high" and "low" value, which differ
by a factor of 5.  The main change made to this Brownian-motion paper is to
adopt these two values and recalculate the resulting integrals that yield the
estimated RMS 10-year uncertainties that are the key result of this paper.
These mass-to-light ratio changes appear as multiplicative scalings of the
previous result.  The thresholds for individually-detectable impulses in Section
3.2.3 have also been updated in concert with the updates to the
mass-determination paper.

Thus the numerical results presented in this paper have been adjusted, but as
per the referee's comments, the methodology remains unchanged.  Changed text is
in red in the draft, and figures that have changed have this noted in their
captions.

A few minor replies to the referee's comments are below.

> This report evaluates the manuscript "Brownian Motion of Main-Belt
> Asteroids on Human Timescales," which examines how gravitational
> interactions between asteroids introduce stochastic positional
> uncertainties over decadal timescales. The study provides a novel
> approach to understanding perturbative effects in asteroid
> ephemerides, with implications for high-precision astronomy.
> 
> The central hypothesis of the study is that the long-term orbital
> evolution of main-belt asteroids is affected by random gravitational
> impulses from other asteroids, leading to a Brownian motion-like
> positional drift. This effect is expected to be particularly
> significant over human timescales (10-100 years), impacting asteroid
> tracking and planetary defense efforts. The study proposes that this
> motion is primarily azimuthal and that current ephemeris models
> underestimate this stochastic contribution.

We did not intend to give the impression that Brownian motion is larger or more
important on decadal scales than over millennia.  Rather our purpose is to
quantify the effect of the Brownian motion on "real-time" dynamics of the solar
system, i.e. the development of ephemerides that fit obervations and adjust
parameters of the solar system's members and of the laws of gravitation.
Understanding the long-term dynamical consequences of the Brownian motion for
population evolution is beyond our scope.

> 
> The study is based on several key assumptions:
> 
> - Asteroid Encounters as Stochastic Events: The model assumes that
> gravitational encounters between asteroids can be treated as a series
> of independent random impulses.
> 
> - Negligibility of Non-Gravitational Effects: The study assumes that
> radiation forces (e.g., Yarkovsky effect) and thermal recoil forces
> have a minor influence compared to gravitational interactions for
> asteroids larger than ~100 m.
> 
> - Mass Uncertainty from LSST Data: The study uses projected LSST data
> to estimate mass uncertainties, heavily relying on Bernstein et
> al. (2025) for these calculations.
> 
> - Choice of Geometric Albedo: The author chooses the geometric albedo
> value that might, in this evaluator's judgment, be too "bright" to be
> representative of the entire population of asteroids in the main
> belt. The problem with that choice lies in the fact that, with that
> value, the mass distribution estimate would be misrepresented. A value
> of 0.15 might be more representative of the average, so I recommend
> its use.
>

See the note at top for a major revision to our estimation of asteroid masses
from their absolute magnitudes.

> In terms of methodology, the study employs a combination of
> theoretical modeling, statistical analysis, and observational
> constraints:
> 
> -Impulse-Based Orbital Perturbation Model: The author developed a
> model where asteroid encounters introduce small, random velocity
> perturbations that accumulate over time, following a Brownian motion
> framework.
> 
> -Empirical Scaling Relations: The study derives scaling relations for
> the expected positional variance over time.
> 
> -Comparison with Gaia and LSST Predictions: The paper estimates how
> well future astronomical surveys can track and mitigate these
> stochastic uncertainties.  Both the mathematical approach and its
> development are appropriate, and I have not detected any errors in
> them.
> 
> Regarding the results, the manuscript highlights some points that
> might be very interesting (the reasons for using "might" are explained
> below):
> 
> - the RMS azimuthal shift due to Brownian motion reaches ~2 km over 10
> years (~1 mas in angular uncertainty).
> 
> - Radial and vertical displacements are found to be significantly
> smaller, with the azimuthal motion dominating the effect.
> 
> - LSST mass uncertainty reductions could improve orbit predictions but
> do not eliminate Brownian motion noise.
> 
> Although the study presents an interesting and underexplored
> perspective on stochastic effects affecting asteroid motion and has a
> robust framework with an application of "Brownian motion" to celestial
> mechanics, it is fully justified to rely on a paper that is in
> preparation (Bernstein et al., 2025) and has clearly not gone through
> the peer review process. This is the reason for using "might" in the
> results section, making it difficult to analyze the final results. If
> the author can improve on this point, the paper would be an
> interesting contribution to the field.
>